---
title: "Fintech Lead Data Analyst SQL Test"
author: "Calvin de Wilde (Sidjaya)"
date: "August 2020"
output: "github_document"
---


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(readr)
library(sqldf)

options(scipen=999)

```

## The Tables

Table name: members

Description: Member ID and information. Contains information for both investor and borrower, which is flagged in the member_type column


```{r, include =FALSE}
members <- read_delim("https://raw.githubusercontent.com/calvindw/dsjobinterviews/main/fintech/members.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)


```

Primary_key = member_id

```{r}
members
```

Table name: loans

Description: Shows the loans and corresponding borrower (member_id) who took the loan. Each loan can only have 1 borrower

```{r, include=FALSE}
loans <- read_delim("https://raw.githubusercontent.com/calvindw/dsjobinterviews/main/fintech/loans.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)
```

primary_key = loan_id

foreign_key = member_id >> members.member_id

```{r}
loans
```

Table name: investments

Description: Shows loan details along with the investors that invested in the loan. Summing the amount will give you the total amount raised for each loan.

foreign_key = member_id >> members.member_id

foreign_key = loan_id >> loans.loan_id


```{r, include=FALSE}
investments <- read_delim("https://raw.githubusercontent.com/calvindw/dsjobinterviews/main/fintech/investments.csv", 
    ";", escape_double = FALSE, col_types = cols(LOAN_ID = col_character(), 
        MEMBER_ID = col_character(), INVESTMENT_DATE = col_datetime(format = "%d-%m-%Y %H:%M")), 
    trim_ws = TRUE)

investments$INVESTMENT_DATE <- as.character(investments$INVESTMENT_DATE)


```


```{r}
investments
```
## SQL Questions

### 1. For each available loan_id, get the corresponding company name.

```{r}
sqldf(" 
SELECT LOAN_ID, COMPANY_NAME  
FROM loans LEFT JOIN members
ON loans.MEMBER_ID = members.MEMBER_ID")


```

### 2. Find out how many borrowers we have per country
```{r}
sqldf("
SELECT COUNTRY_CODE, count(MEMBER_TYPE) AS n_count
FROM members
WHERE (MEMBER_TYPE='borrower')
GROUP BY COUNTRY_CODE"
)

```

### 3. Find out how many investors per CITIZENSHIP_COUNTRY invested in loan 9391

```{r}

sqldf("
select 
members.CITIZENSHIP_COUNTRY,
count(members.CITIZENSHIP_COUNTRY) as investors_count
from loans
left join members 
        on members.MEMBER_ID = loans.MEMBER_ID
where loans.LOAN_ID = '9391'
group by members.CITIZENSHIP_COUNTRY
      ")
```


### 4. Find out the total amount invested and number of investors who participated in at least loan in each month per year, for loans originating in SG (i.e. loans by SG borrowers)

```{r}

sqldf("
with 
  members_filtered AS(
                      select MEMBER_ID,CITIZENSHIP_COUNTRY
                      from members
                      where (members.MEMBER_TYPE = 'borrower' AND members.COUNTRY_CODE = 'SG')
                      )

select 
  strftime('%Y', INVESTMENT_DATE) as year,
  strftime('%m', INVESTMENT_DATE) as month,
  count(members_filtered.MEMBER_ID) as n_investor
from members_filtered
left join  investments 
    on investments.MEMBER_ID = members_filtered.MEMBER_ID
group by year, month
having n_investor >= 1
      ")

```

### 5. Find out which loan(s) raised more than USD10,000 in 2018.


```{r}
sqldf("
      select 
        LOAN_ID,
        sum(investments.'AMOUNT (USD)') as sum_amount
      from investments
      where (INVESTMENT_DATE >= '2018-01-01' and 
              INVESTMENT_DATE <= '2018-12-31')
      group by LOAN_ID
      having sum_amount > 10000
      ")


```

alternative version, using between


```{r}
sqldf("
      select 
        LOAN_ID,
        sum(investments.'AMOUNT (USD)') as sum_amount
      from investments
      where INVESTMENT_DATE between '2018-01-01' and '2018-12-31'
      group by LOAN_ID
      having sum_amount > 10000
      ")


```

### 6. Find the top 10 largest loans in terms of amount raised in 2017 per country. Do not use UNION as we do not want to query the table multiple times.


```{r}
#you can use where between to filter the date too, see solution   #5

sqldf("
with  cte as (select 
                LOAN_ID,
                COUNTRY_CODE,
                sum(investments.'AMOUNT (USD)') as sum_amount
              from investments
              left join members
                on investments.MEMBER_ID=members.MEMBER_ID
              where (INVESTMENT_DATE >= '2017-01-01' AND 
                     INVESTMENT_DATE <= '2017-12-31')
              group by LOAN_ID, COUNTRY_CODE)
              

select 
COUNTRY_CODE,
LOAN_ID,
rank () over (partition by COUNTRY_CODE order by -sum_amount) as loan_ranking_by_country,
sum_amount
from cte
              
      ")


```


## Oral questions

After passing the first round interview I was interviewed by the company's lead analyst related to my knowledge on the business acumen. I thought I was doing a good job until he suddenly asked me to answer another SQL questions. There were 3 of them, but I could only share two as the last one was doing a live full join. 

### 1.What are the different types of joins? 

left, right, inner, full, cross join. I forgot cross join since I never used it.

### 2. What's the difference between RANK, DENSE_RANK, and ROW_NUMBER?

this [answer](https://stackoverflow.com/questions/11183572/whats-the-difference-between-rank-and-dense-rank-functions-in-oracle) and this [answer](https://stackoverflow.com/questions/7747327/sql-rank-versus-row-number) from SO give a good definition:

RANK: RANK gives you the ranking within your ordered partition. Ties are assigned the same rank, with the next ranking(s) skipped. So, if you have 3 items at rank 2, the next rank listed would be ranked 5

DENSE_RANK again gives you the ranking within your ordered partition, but the ranks are consecutive. No ranks are skipped if there are ranks with multiple items.

ROW_NUMBER: Returns a unique number for each row starting with 1. For rows that have duplicate values,numbers are arbitarily assigned.


### 3. Which join I need to use when I want to show the sum of transactions between two tables? Let's say table for transactions in January and February, both of them have unique user_id and transaction column.

For this situation, use full join, followed by group by and select statement